---
title: "LASSO"
format: 
  html:
    html-math-method: katex

engine: knitr
filters:
- webr

webr:
  show-startup-message: false      # Display status of webR initialization
  show-header-message: false      # Check to see if COOP&COEP headers are set for speed.
#  packages: ['ggplot2', 'dplyr'] # Pre-install dependencies
#  home-dir: "/home/rstudio"      # Customize where the working directory is
#  base-url: ''                   # Base URL used for downloading R WebAssembly binaries
#  service-worker-url: ''         # URL from where to load JavaScript worker scripts when loading webR with the ServiceWorker communication channel.
---

<style>
    main {
        text-align: justify;    
    }
</style>
<script src="https://cdn.plot.ly/plotly-latest.min.js"></script>

**Prerequisites:** <br>

For this tutorial, we assume that you have a basic knowledge of linear regression, especifically the least squares method and inference for the regression coefficients; and that you  have been introduced to cross-validation, generalization error and overfitting/underfitting concepts. 
    

**Learning Objectives:** <br>

At end of this tutorial you should be able to:

- Describe the Lasso method and its advantages over the least squares;
- Constrast Lasso with the least squares;
- Recognize the impact the Lasso method has on the inference of the regression coefficients;
- Identify the problem of applying Lasso with categorical variables;
- Apply Lasso and Group Lasso to a dataset using R;

<hr>

## Prostate Cancer Data 

Let us take a problem-driven approach to introduce LASSO. Let's start by loading some packages we will use. 

```{r warning=FALSE, message=FALSE}
# Loading some libraries
library(tidyverse) 
library(tidymodels)
library(knitr)
library(broom)
library(leaps)
library(glmnet)
library(Brq) # This package contains the data
```

We will use a dataset of 97 patients with prostate cancer (Stamey et al. 1989). The dataset is available in the `Brq` R package. Before starting the analysis, we should understand what data we have. The following are the variables in our data frame:

- **lcavol**: log(cancer volume) 
- **lweight**: log(prostate weight)
- **age**: age of the patient
- **lbph**: log(amount of benign prostatic hyperplasia)
- **svi**: Seminal vesicle invasion. (True or False)
- **lcp**: log(capsular penetration)
- **gleason**: Gleason score. The Gleason score measures how abnormal the tissue looks. The lower the score is, the more the cells look like regular prostate tissue. You can learn more about the [Gleason Score here](https://www.pcf.org/about-prostate-cancer/diagnosis-staging-prostate-cancer/gleason-score-isup-grade/)
- **pgg45**: Percentage Gleason scores 4 or 5. Scores 4 and 5 are considered very high; the cells barely look like normal prostate tissue.
- **lpsa**: log(prostate-specific antigen)

Unfortunately, the units for the variables are not easily available and are unknown. However, keep in mind that the units are fundamental to interpreting the model. So always make sure you know what the units are (even though I'm not sure what the units are in this case &#x1F609;). 

Let's load the data and take a look at the first few rows.

```{r warning=FALSE, message=FALSE}
# Loading the data
data("Prostate")
cancer_prostate <- tibble(Prostate)

# Printing the first 10 rows
kable(head(cancer_prostate, 10), digits = 4)
```
<br>
The first thing we need to do is to split the data into a training and a test set. We will use the training set to fit the model, including variable selection and the test set to evaluate the model's performance. So let's randomly split our data into 70% training and 30% test sets.

```{r}
set.seed(1) # Let's set the seed for reproducibility

cancer_split <- initial_split(cancer_prostate, prop = 0.7)
cancer_train <- training(cancer_split)
cancer_test <- testing(cancer_split)
```

Now that we have our training set, we can use it how we want, as much as we want. Let's start simple and look at some summary statistics of our variables. We will remove `svi` because it is a dichotomic variable (true or false) with a proportion of true of `r round(mean(cancer_train$svi),3)`.

```{r}
summary(cancer_train |> select(-svi))
```

We want to model `lpsa`, the log(prostate-specific antigen), using the remaining variables as explanatory variables (or independent variables, or predictors, or features, or covariates -- so many names for the same things, I know!). There are a wide variety of mathematical models we could use, linear Regression being one of the simplest ones. However, Linear Regression has the advantage of being highly interpretable while performing very well in many prediction tasks. 

## Linear Regression

In regression, we want to model the relationship between a numerical variable $Y$ and some, say $p$, explanatory variables $X_1, X_2, \ldots, X_p$. The linear regression equation is given by: 

$$
Y_i = \beta_0 + \beta_1X_{1i} + \beta_2X_{2i} + \ldots + \beta_3X_{pi} + \varepsilon_i,\quad i=1,\ldots,n
$$
where $\varepsilon$ is random error. 

In the Cancer Prostate case, our linear regression model is given by:

\begin{align*}
\underbrace{lpsa_i}_\text{$Y_i$} = &\ \beta_0 + \beta_1\times \underbrace{lcavol_i}_\text{$X_{i1}$} + \beta_2\times \underbrace{lweight_i}_\text{$X_{i2}$} + \beta_3\times \underbrace{age_i}_\text{$X_{i3}$} + \beta_4\times \underbrace{lbph_i}_\text{$X_{i4}$}\ + \\
& + \beta_5\times \underbrace{svi_i}_\text{$X_{i5}$} + \beta_6\times \underbrace{lcp_i}_\text{$X_{i6}$} + \beta_7\times  \underbrace{gleason_i}_\text{$X_{i7}$} + \beta_8\times \underbrace{pgg45_i}_\text{$X_{i8}$} + \varepsilon_i
\end{align*}

To fit this model to the data at hand, we find the values for the coefficients $\bm{\beta} = \left(\beta_0, \beta_1, \ldots, \beta_p\right)$ that minimizes the *residual sum of squares* (RSS), i.e., 

$$
    RSS(\bm{\beta}) = \sum_{i=1}^n \left(y_i - \beta_0 - \beta_1X_{i1} - \ldots - \beta_p X_{ip}\right)^2
$$

```{r}
# Fitting the linear regression model
cancer_lm <- lm(lpsa ~ ., data = cancer_train)

# Printing the coefficients' information
kable(tidy(cancer_lm))
```

So the values in the estimate columns are the values that minimize `RSS` for our training data of the Prostate Cancer dataset. If you are not very familiar with statistical inference, you can ignore the other columns.

### Improving the Ordinary Least Square (OLS)

Basically, there are two reasons to fit a linear regression model: 

- **Prediction**: We want to predict the response based on the values of the predictors the best we can. So, we want our model to generalize well to unseen data. 

- **Inference**: We are interested in interpreting the association between our response and the predictors and checking their significance. 

For prediction, we want to avoid overfitting to training the data. Remember, overfitting occurs when your model does very well in the training data but poorly in unseen data. A common approach to avoid overfitting is the so-called *Regularization*. Regularization means we restrict (or penalize) our optimization a little bit by adding a penalty term to our cost function. 

For inference, we should use the simplest model possible. What do we mean by the simplest model here? The model with the fewest predictors. There are mainly two reasons for this: (1) it is easier to interpret a model with fewer covariates; and (2) the principle of parsimony (Occam's razor). As you can imagine, selecting only the relevant covariates for our model is vital in statistical modelling. 

We shall now discuss both aspects: (1) variable selection and (2) regularization. As it turns out, the LASSO method we will discuss next is a great tool for both purposes.


#### Variable Selection

The easiest way (intuitively, at least) to improve our model is to try the model with all possible combinations of covariates, and then we select the best combination for our final model. Simple, eh? Yeah, the problem is that, even for a relatively small number of covariates, it quickly becomes impossible to check all possible models. The table below shows the number of models we need to try for different numbers of covariates. 

|Number of Covariates   |Number of Models to Fit  |
|-----------------------|-------------------------|
|5                      |31                       |
|10                     |1023                     |
|20                     |1,048,575                |
|30                     |1,073,741,823            |
|40                     |1,099,511,627,775        |
|50                     |1,125,899,906,842,623    |

Can you imagine fitting over 1 trillion models to find the best subset of covariates to use? Because of this, people have developed some strategies, such as *Backward Selection*, *Forward Selection*. 

In **Backward Selection**, we first consider all the covariates. Then, we look for the least important covariate and remove it from the model. We do this iteratively until we have the best model. The **Forward Selection** works in a similar fashion, but we start with the null model (no covariates) and add one covariate at a time. There are mainly two problems with using Backward or Forward selection. 

The first problem is that these methods are not guaranteed to find the best possible subset of covariates. Instead of trying 1,099,511,627,775 for 40 covariates, we will try 820 models with Backward or Forward selection. Much better, but plenty of possible combinations of covariates are not being considered. 

The second problem is that this is not a smooth process. We either include a covariable or not. There is no "partially" including a variable. This lack of smoothness makes the results of these methods quite sensitive to small numerical changes in the data. Then it comes regularization. 

#### Regularization

As we know, to fit a linear regression we look for the coefficients $\bm{\beta}$ that minimizes the Residual Sum of Square (RSS). So any $\bm{\beta}$ is fair game. We are only interested in driving our RSS down. 

The idea of regularization is to restrict our search a bit by imposing a cap on our $\bm{\beta}$. So basically, we are saying:

> ok, I'm looking to minimize the RSS, but I don't want my $\bm{\beta}$ to get too crazy (in other words, very high or very low values). 

But how do we do this? We will add a penalty term to our RSS to tell our model that there is a price to be paid to use a high value (positive or negative) for $\bm{\beta}$:

$$
\underbrace{\sum_{i=1}^n \left(y_i - \beta_0 - \beta_1X_{i1} - \ldots - \beta_p X_{ip}\right)^2}_\text{RSS} + {\color{red} \text{penalty term}}
$$

There are many penalties we can add to our RSS. Two widely used ones are:

- $\lambda\sum_{i=1}^p \beta_i ^2$ this leads to the so-called **Ridge Regression**. 

- $\lambda\sum_{i=1}^p \left|\beta_i\right|$ this leads to the so-called **LASSO**.

**Important:** Note that the we do not include the intercept $\beta_0$. 

The penalty term has two components a positive constant $\lambda$ and a norm of $\bm{\beta}$. It is like when you go to the grocery store to buy potatoes. The constant $\lambda$ is the price of the potatoes, and the norm (either $\sum_{i=1}^p \beta_i ^2$ or $\sum_{i=1}^p \left|\beta_i\right|$) measures the amount of potatoes you are buying. The different norms refer to different ways to measure the amount of potatoes being bought. For example, the grocery store could charge you per pound of potatoes, or it could charge you per unit of potatoes. 

Naturally, the higher the price of the potato, the fewer potatoes you will buy. The same happens with the constant $\lambda$. The higher $\lambda$ is, the smaller the values for $\bm{\beta}$ we will use. Note that increasing the magnitude of any $\beta_j$ will increase the penalty term, and it will only be acceptable if there is a significant decrease in the RSS to compensate for the increase in the penalty term. Because of the forced reduction these penalties cause in the magnitude of the coefficients $\bm{\beta}$, they are frequently called  **Shrinkage** methods. 

Now, the penalty term ends up "measuring the size" of the coefficients. Look at the summary quantities of our variables above. Compare the magnitude of the predictors with the response. The predictors are in different scales. So, if we add some penalties according to the size of the coefficients, we need to put all predictors on the same scale. Otherwise, some variables will be more penalized than others. We do this by **standardizing the predictors**, a step that **must** be done in regularization. 

Regularization methods help us to mitigate overfitting and, in the case of LASSO, it helps us with variable selection. 

##### Why shrinking $\beta$ regularizes the model?



In the following, we will restrict our focus to LASSO. 


## The LASSO method


The LASSO method, which stands for Least Absolute Shrinkage and Selection Operator, was proposed by Tibshirani (1989). It adds an L1-norm regularization penalty to the RSS function that we minimize when fitting the Ordinary Least Square Regression. After the inclusion of the penalty term, the function we try to minimize is given by:

$$
LASSO(\bm{\beta}, \lambda) = \sum_{i=1}^n \left(y_i - \beta_0 - \beta_1X_{i1} - \ldots - \beta_p X_{ip}\right)^2 + {\color{red} \lambda\sum_{i=1}^p \left|\beta_i\right|}
$$

Another way we could frame the LASSO method is the following. We want to minimize the RSS (just like the OLS) but subject to the condition that the sum of the absolute values of coefficients is less than a positive number. Mathematically, 

$$
\text{minimize } \sum_{i=1}^n \left(y_i - \beta_0 - \beta_1X_{i1} - \ldots - \beta_p X_{ip}\right)^2 
$$
subject to
$$
\sum_{i=1}^p \left|\beta_i\right| \leq C, \text{ for some } C>0.
$$

This second formulation of LASSO is saying that we want to minimize the RSS (just like in the OLS). But differently than the OLS, not all the $\bm{\beta}$ are fair game. We need a $\bm{\beta}$ that satisfies the condition $\sum_{i=1}^p \left|\beta_i\right| \leq C$ for a positive constant $C$. Doing this constrains the search space of the solution: we don't look for the best $\bm{\beta}$ everywhere; we look for the best $\bm{\beta}$ in a given smaller region. 

### Example 1: LASSO with two covariates

To be able to visualize what the LASSO method is doing, let's restrict our attention to a small case of two covariates. Although this is of little practical interest, it will be useful for us to be able to make some plots. 

Let us consider only two covariates of the Prostate Cancer dataset: `lcavol` and `lweight`. Let's first standardize the predictors. In this case we will also center the response, just to get rid of the intercept (as it will be 0). 

```{r}
# Standardizing the predictors and centering the response
cancer_train_std <- 
    cancer_train |>
    select(lcavol, lweight, lpsa) |>
    mutate(lcavol = (lcavol - mean(lcavol)) / sd(lcavol),
           lweight = (lweight - mean(lweight)) / sd(lweight),
           lpsa = (lpsa - mean(lpsa)))
```

Let's double check that everything is in order (for our mental sanity).

```{r}
cancer_train_std |>
    summarise(across(everything(), list(mean = mean, sd = sd))) |>
    pivot_longer(cols = everything(), names_to = c("variable", '.value'), names_sep = "_") |>
    kable(digits = 4)
```

Ok, everything looks good. All variables are centered (i.e., have 0 mean) and only the predictors are scaled (i.e., have std. deviation of 1). Our regression model is: 

$$
lpsa_i = \beta_1\times lcavol_i + \beta_2\times lweight_i + \varepsilon_i
$$
Note that there's no intercept. We already removed it by centring all the variables. So then, all that is left for us to do is estimate $\beta_1$ and $\beta_2$.


#### The Ordinary Least Square (OLS) Solution

The OLS estimates for $\beta_1$ and $\beta_2$ are the ones that minimize the RSS, shown in the figure below. The green point is the OLS solution. 

```{r, include=FALSE}
model_two_vars <- lm(lpsa ~ 0 + lcavol + lweight, data = cancer_train_std)

coef_lpsa_two_vars <- 
    model_two_vars  |>
    coef() |>
    as.numeric()

rss_two_vars <- sum(model_two_vars$residuals**2)
rss_two_vars
```

```{python, echo=FALSE}
#| label: RSS
import plotly.graph_objects as go
import pandas as pd
import numpy as np

# Read data from R
cancer_train_std_py = r.cancer_train_std[['lcavol', 'lweight', 'lpsa']]


# Create a Meshgrid for plotting
x = np.linspace(-4, 4, 100)
y = np.linspace(-4, 4, 100)
xx, yy = np.meshgrid(x, y)

# Calculate the RSS for the meshgrid
z_data = pd.DataFrame({
    'b1': np.ravel(xx),
    'b2': np.ravel(yy),
    'data': np.array(
        [np.sum((cancer_train_std_py['lpsa'].to_numpy() - b1*cancer_train_std_py['lcavol'].to_numpy() - b2*cancer_train_std_py['lweight'].to_numpy())**2) 
         for b1, b2 in zip(np.ravel(xx), np.ravel(yy))])}
)


fig = go.Figure(data=[go.Surface(x = x, y = y, z=z_data['data'].to_numpy().reshape(len(x), len(y)))])

fig.update_layout(
    title={
        'text': 'Residual Sum of Square as a function of the coefficients',
        'y': 0.9,	
        'x': 0.45,
        'xanchor': 'center',
        'yanchor': 'top'},
    autosize=False,
    width=800,
    height=500,
    margin=dict(l=15, r=10, b=60, t=40),
    scene=dict(
        xaxis_title=r'β1',
        yaxis_title=r'β2',
        zaxis_title=r'RSS',
    )).add_trace(go.Scatter3d(
        x = [r.coef_lpsa_two_vars[0]], 
        y = [r.coef_lpsa_two_vars[1]], 
        z = [r.rss_two_vars], 
        mode = "markers", type = "scatter3d", 
        marker = {'size': 15, 'color': "green", 'symbol': 'circle'}))
    #yaxis_title='b2',
    #zaxis_title='RSS')
```
<br>

```{r}
# Fitting the OLS model with no intercept.
ols_lm <- lm(lpsa ~ 0 + ., data = cancer_train_std)
kable(tidy(ols_lm), digits = 4)
```

The OLS solutions is given by $\beta_1 = `r round(coef_lpsa_two_vars[1], 4)`$ and $\beta_2 = `r round(coef_lpsa_two_vars[2], 4)`$. The RSS at this point is `r round(rss_two_vars, 4)`. 

#### The LASSO Solution

If we decide to apply Lasso, with a constraint of $\left|\beta_1\right| + \left|\beta_2\right| \leq 0.5$ (this is equivalent to using $\lambda = 0.311$), the OLS solution would not be a valid solution because $`r round(coef_lpsa_two_vars[1], 4)` + `r round(coef_lpsa_two_vars[2], 4)` = `r round(coef_lpsa_two_vars[1], 4) + round(coef_lpsa_two_vars[2], 4)`\geq 0.5$. The only solutions that would be acceptable would be the ones that lie within the square shown in the figure below. 

```{r echo = FALSE}
ggplot(tibble(x1 = -0.5,  x2 = 0, x3 = 0.5, x4 = 0,
              y1 = 0, y2 = 0.5, y3 = 0, y4 = -0.5)) + 
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2)) +
  geom_segment(aes(x = x2, y = y2, xend = x3, yend = y3)) +
  geom_segment(aes(x = x3, y = y3, xend = x4, yend = y4)) +
  geom_segment(aes(x = x1, y = y1, xend = x4, yend = y4)) +
  labs(x = 'β1', y = 'β2') +
  xlim(-1, 1) +
  ylim(-1, 1) +
  geom_point(aes(x,y), color = "darkgreen", size = 3, data = tibble(x=coef_lpsa_two_vars[1], y=coef_lpsa_two_vars[2])) + 
  annotate("text", x = coef_lpsa_two_vars[1]-0.02, y = coef_lpsa_two_vars[2] + 0.06, label = "OLS solution") + 
  theme(plot.background = element_rect(fill="#ffffff"),
        panel.background = element_rect(fill="#ffffff"),
        text = element_text(size=14),
        panel.grid.major=element_line(colour="#00000014"),
        panel.grid.minor=element_line(colour="#00000014"))
```

The LASSO method then needs to shrink the coefficients until they are in the square. So, the LASSO solution is the best solution in the feasible region (the square).

```{r, include=FALSE}
fit <- glmnet(cancer_train_std |> select(lcavol, lweight), cancer_train_std$lpsa, lambda = 0.311)
kable(tidy(fit))
```

```{r echo = FALSE}
lasso_coef = coef(fit, s=0.311, exact = TRUE)
ggplot(tibble(x1 = -0.5,  x2 = 0, x3 = 0.5, x4 = 0,
              y1 = 0, y2 = 0.5, y3 = 0, y4 = -0.5)) + 
  geom_segment(aes(x = x1, y = y1, xend = x2, yend = y2)) +
  geom_segment(aes(x = x2, y = y2, xend = x3, yend = y3)) +
  geom_segment(aes(x = x3, y = y3, xend = x4, yend = y4)) +
  geom_segment(aes(x = x1, y = y1, xend = x4, yend = y4)) +
  labs(x = 'β1', y = 'β2') +
  xlim(-1, 1) +
  ylim(-1, 1) +
  geom_point(aes(x,y), color = "darkgreen", size = 3, data = tibble(x=coef_lpsa_two_vars[1], y=coef_lpsa_two_vars[2])) + 
  annotate("text", x = coef_lpsa_two_vars[1]-0.02, y = coef_lpsa_two_vars[2] + 0.06, label = "OLS solution") + 
  geom_point(aes(x,y), color = "red", size = 3, data = tibble(x=lasso_coef[2], y=lasso_coef[3])) + 
  annotate("text", x = lasso_coef[2], y = lasso_coef[3] - 0.06, label = "LASSO solution") +
  theme(plot.background = element_rect(fill="#ffffff"),
        panel.background = element_rect(fill="#ffffff"),
        text = element_text(size=14),
        panel.grid.major=element_line(colour="#00000014"),
        panel.grid.minor=element_line(colour="#00000014"))
```

As we can see from the figure above, the LASSO solution was in a vertice of the square, where one of the coefficients is zero. This is not unusual. The LASSO solution will generally be in a vertice, where one or more of the coefficients are zero. For this reason, LASSO is also considered a variable selection method. The LASSO coefficients, in this case, were $\beta_1 = 0.5$ and $\beta_2=0$. 

We could also look at this on the perspective of the loss function $LASSO(\bm\beta, \lambda)$. Naturally, this function is a function of $\lambda$. The figure below shows the loss function for different values of $\lambda$. 

<div id='rss-plot-lambdas'>

</div>

```{r}
#| output: false
#| echo: false
fit_many_lambdas <- glmnet(cancer_train_std |> select(lcavol, lweight), cancer_train_std$lpsa, lambda = seq(0, 4000, 0.05))

coef_matrix <- coef(fit_many_lambdas) |> as.matrix()
ojs_define(coef_matrix = coef_matrix)
ojs_define(cancer_data = cancer_train_std |> select(lpsa, lcavol, lweight) |> as.matrix())
```

```{ojs}
//| echo: false
// Create a Meshgrid for plotting
values_plot = {
    let b1 = Array(100).fill().map((element, index) => -4 + index*0.08);
    let b2 = Array(100).fill().map((element, index) => -4 + index*0.08);
    

    let rss = [];
    for (let i = 0; i < b1.length; i++) { 
        rss[i] = [];
        for (let j = 0; j < b2.length; j++) { 
            rss[i].push(cancer_data.map(elem => (elem[0] - b1[i] * elem[1] - b2[j] * elem[2]) ** 2 + lambda * (Math.abs(b1[i]) + Math.abs(b2[j])) ).reduce((a, t) => a + t, 0));
        }
    };

    return {
        b1: b1,
        b2: b2,
        rss: rss
    };
};

```

```{ojs} 
//| echo: false
//| include: true

viewof lambda = Inputs.range([0, 3], {value: 0, step: 0.01, label: "Lambda"});

garbage = {
let surfaceData = {
    type: 'surface',
    x: values_plot.b1,
    y: values_plot.b2,
    z: values_plot.rss
};

let layout = {
    title: 'RSS Surface Plot',
    scene: {
        xaxis: { title: 'Beta 1' },
        yaxis: { title: 'Beta 2' },
        zaxis: { title: 'RSS' }
    }
};

Plotly.newPlot('rss-plot-lambdas', [surfaceData], layout);
};
```

- What happens when you increase $\lambda$ to very large values?
- What happens when you decrease $\lambda$ to zero? 
- Once we start increasing $\lambda$, all of our $\beta$ coefficients will never increase. 

### Group Lasso



---

```{python, echo=FALSE}
#| label: LASSO
#| include: false
import plotly.graph_objects as go
import pandas as pd
import numpy as np

# Read data from R
cancer_train_std_py = r.cancer_train_std[['lcavol', 'lweight', 'lpsa']]


# Create a Meshgrid for plotting
x = np.linspace(-5, 5, 100)
y = np.linspace(-5, 5, 100)
xx, yy = np.meshgrid(x, y)

# Calculate the RSS for the meshgrid
z_data = pd.DataFrame({
    'b1': np.ravel(xx),
    'b2': np.ravel(yy),
    'data': np.array(
        [np.sum((cancer_train_std_py['lpsa'].to_numpy() - b1*cancer_train_std_py['lcavol'].to_numpy() - b2*cancer_train_std_py['lweight'].to_numpy())**2) + 1000*0.311*(abs(b1) + abs(b2)) 
         for b1, b2 in zip(np.ravel(xx), np.ravel(yy))])}
)

#z_data

fig = go.Figure(data=[go.Surface(x = x, y = y, z=z_data['data'].to_numpy().reshape(len(x), len(y)))])

fig.update_layout(
    title={
        'text': 'Residual Sum of Square as a function of the coefficients',
        'y': 0.9,	
        'x': 0.45,
        'xanchor': 'center',
        'yanchor': 'top'},
    autosize=False,
    width=800,
    height=500,
    margin=dict(l=15, r=10, b=60, t=40),
    scene=dict(
        xaxis_title=r'β1',
        yaxis_title=r'β2',
        zaxis_title=r'RSS',
    )).add_trace(go.Scatter3d(
        x = [0.5], 
        y = [0], 
        z = [r.rss_two_vars], 
        mode = "markers", type = "scatter3d", 
        marker = {'size': 15, 'color': "green", 'symbol': 'circle'}))
    #yaxis_title='b2',
    #zaxis_title='RSS')
```