---
title: "LASSO"
format: html
engine: knitr
#webr:
#  show-startup-message: false      # Display status of webR initialization
#  show-header-message: false      # Check to see if COOP&COEP headers are set for speed.
#  packages: ['ggplot2', 'dplyr'] # Pre-install dependencies
#  home-dir: "/home/rstudio"      # Customize where the working directory is
#  base-url: ''                   # Base URL used for downloading R WebAssembly binaries
#  service-worker-url: ''         # URL from where to load JavaScript worker scripts when loading webR with the ServiceWorker communication channel.
filters:
- webr
---

<style>
    main {
        text-align: justify;    
    }
</style>


**Prerequisites:** <br>

For this tutorial, we assume that you have a basic knowledge of linear regression, especifically the least squares method and inference for the regression coefficients; and that you  have been introduced to cross-validation, generalization, error and overfitting/underfitting concepts. 
    

**Learning Objectives:** <br>

At end of this tutorial you should be able to:

- Describe the Lasso method and its advantages over the least squares;
- Constrast Lasso with the least squares;
- Recognize the impact the Lasso method has on the inference of the regression coefficients;
- Identify the problem of applying Lasso with categorical variables;
- Apply Lasso and Group Lasso to a dataset using R;

<hr>

### Linear regression

Let us take a problem-driven approach to introduce LASSO. Let's start by loading some packages we will use. 

```{r warning=FALSE, message=FALSE}
# Loading some libraries
library(tidyverse) 
library(tidymodels)
library(knitr)
library(Brq) # This package contains the data
```

We will use a dataset of 97 patients with prostate cancer (Stamey et al. 1989). The dataset is available in the `Brq` R package. Before starting the analysis, we should understand what data we have. The following are the variables in our data frame:

- **lcavol**: log(cancer volume) 
- **lweight**: log(prostate weight)
- **age**: age of the patient
- **lbph**: log(amount of benign prostatic hyperplasia)
- **svi**: Seminal vesicle invasion. (True or False)
- **lcp**: log(capsular penetration)
- **gleason**: Gleason score. The Gleason score measures how abnormal the tissue looks. The lower the score is, the more the cells look like regular prostate tissue. You can learn more about the [Gleason Score here](https://www.pcf.org/about-prostate-cancer/diagnosis-staging-prostate-cancer/gleason-score-isup-grade/)
- **pgg45**: Percentage Gleason scores 4 or 5. Scores 4 and 5 are considered very high; the cells barely look like normal prostate tissue.
- **lpsa**: log(prostate-specific antigen)

Unfortunately, the units for the variables are not easily available and are unknown. However, keep in mind that the units are fundamental to interpreting the model. So always make sure you know what the units are (even though I'm not sure what the units are in this case &#x1F609;). 

Let's load the data and take a look at the first few rows.

```{r warning=FALSE, message=FALSE}
# Loading the data
data("Prostate")
cancer_prostate <- tibble(Prostate)

# Printing the first 10 rows
kable(head(cancer_prostate, 10))
```

```{webr-r}
# Hello
```

The first thing we need to do is to split the data into a training and a test set. We will use the training set to fit the model, including variable selection, and the test set to evaluate the model's performance. 

```{r}
cancer_split <- initial_split(cancer_prostate, prop = 0.7)
cancer_train <- training(cancer_split)
cancer_test <- testing(cancer_split)
```

```{r}
summary(cancer_train)
```



<!-- 
```{r}
library(aplore3)
birth_weight <- 
    lowbwt |>
    rename(
        age = age,
        weight_lbs = lwt,
        race = race,
        had_premature_labor = ptl,
        n_visit_first_trimester = ftv,
        uterine_irritability = ui,
        baby_birth_weight_g = bwt) |> 
    select(-low)
```
A baby's weight at the moment of birth is an important indication of the baby's health. For example, there have been some associations between high birth weight and increased risk of developing certain types of cancer (e.g., High Birth Weight Increases the Risk for Bone Tumor: A Systematic Review and Meta-Analysis - PubMed (nih.gov)) and low birth weight might increase the risk of childhood asthma (Effect of low birth weight on childhood asthma: a meta-analysis - PubMed (nih.gov)) among other complications. Below is a dataset about 189 mothers and their babies weight at birth. 

```{r}
knitr::kable(head(birth_weight))
```
-->



In regression, we want to model the relationship between a numerical variable $Y$ and some, say $p$, explanatory variables $X_1, X_2, \ldots, X_p$. There are different ways we can model such a relationship. Linear regression is one of the simplest (although still quite powerful) models. The linear regression is given by: 

$$
Y_i = \beta_0 + \beta_1X_{1i} + \beta_2X_{2i} + \ldots + \beta_3X_{pi} + \varepsilon,\quad i=1,\ldots,n
$$


```{python}

import plotly.graph_objects as go
import pandas as pd

# Read data from a csv
z_data = pd.read_csv('https://raw.githubusercontent.com/plotly/datasets/master/api_docs/mt_bruno_elevation.csv')

fig = go.Figure(data=[go.Surface(z=z_data.values)])

fig.update_layout(title='Mt Bruno Elevation', autosize=False,
                  width=800, height=500,
                  margin=dict(l=65, r=50, b=65, t=90))

fig.show()
```