---
bibliography: ../references.bib
format: 
  revealjs:
    margin: 0.05
    max-scale: 4
    chalkboard: true
    theme: [default, ../style.scss, lecture02.scss]
    slide-number: c
    show-slide-number: all
    preview-links: auto
    footer: "&copy; 2023 Rodolfo Lourenzutti & Eugenia Yu -- Material Licensed under [CC By-SA 4.0](https://creativecommons.org/licenses/by-sa/4.0/)"
---



# Exploratory Data Analysis: <br> Categorical Data {background-image="imgs/caravel-transparent.png" background-size="contain" background-opacity="0.15"}

STAT 200 - Lecture 2

```{r}
library(tidyverse)
```


## Exploratory Data Analysis (EDA)

-   One of the most critical steps when studying a new problem and data set.

-   EDA helps us to:

    -   better understand the data;
    -   find new and frequently unexpected relationships between variables;
    -   raise interesting questions about the study;

## Disclaimer

-   You are in the driver's seat of your EDA.
    -   There's no recipe;
    -   Little thought yields small findings
-   In general, EDA findings are preliminary that require more investigation.

# Categorical Variables

## Data

-   For this section, let's consider the dataset of [schools in Vancouver](https://opendata.vancouver.ca/explore/dataset/schools/information/).

<iframe src="https://opendata.vancouver.ca/explore/embed/dataset/schools/table/?&amp;static=false&amp;datasetcard=false" width="800" height="300" frameborder="0" style="margin-left:12vh; margin-right:5vh; margin-bottom: 0;">
</iframe>

:::{.fragment}
- which of the variables above in this data is categorical? 
:::

## Frequency Tables

::: {style="margin-top: 10vh;"}
:::

```{r}
schools <- read_delim("https://opendata.vancouver.ca/explore/dataset/schools/download/?format=csv&timezone=America/Los_Angeles&lang=en&use_labels_for_header=true&csv_separator=%3B", delim = ';')

frequency_table <- 
  schools |> 
  rename(school_category = SCHOOL_CATEGORY) %>% 
  mutate(school_category = as_factor(school_category)) %>% 
  count(school_category)

frequency_table %>% 
  rename(`School Category` = school_category) |>
  knitr::kable()
```

## Bar Chart

:::: {.columns}

::: {.column width="50%"}

```{r, out.size="1000px"}
#| height: 600px
#| 
frequency_table |>
  ggplot() + 
  geom_bar(aes(fct_reorder(school_category, n, .desc = TRUE), n), stat = "identity") + 
  xlab("") +
  ylab("Frequency") + 
  ggtitle("Number of schools in Vancouver") +
  theme(text = element_text(size=22))
```

:::

::: {.column width="50%"}

```{r}
#| height: 600px
#| 
frequency_table |>
  ggplot() + 
  geom_bar(aes(n, fct_reorder(school_category, n, .desc = TRUE)), stat = "identity") + 
  ylab("") +
  xlab("Frequency") + 
  ggtitle("Number of schools in Vancouver") +
  theme(text = element_text(size=22))

```

:::

::::


## Pie Chart (Avoid!){.smaller}


```{r}
# Compute the position of labels
frequency_table <- 
  frequency_table %>% 
  mutate(prop = n / sum(frequency_table$n) * 100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop )

# Basic piechart
ggplot(frequency_table, aes(x="", y=prop, fill=school_category)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  theme_void() + 
  labs(fill = "School Category") +
  theme(text = element_text(size=20))+
  ggtitle("Number of schools in Vancouver")

  #theme(legend.position="none") +
  #geom_text(aes(y = ypos, label = school_category), color = "white", size=4) +
  #scale_fill_brewer(palette="Set1")
```



## Pie Chart (Avoid!){.smaller}

:::: {.columns}

::: {.column width="50%"}
```{r}

# Compute the position of labels
frequency_table <- 
  frequency_table %>% 
  mutate(prop = n / sum(frequency_table$n) * 100) %>%
  mutate(ypos = cumsum(prop) - 0.5 * prop )

# Basic piechart
ggplot(frequency_table, aes(x="", y=prop, fill=school_category)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  theme_void() + 
  labs(fill = "School Category") +
  theme(text = element_text(size=20))+
  ggtitle("Number of schools in Vancouver")

  #theme(legend.position="none") +
  #geom_text(aes(y = ypos, label = school_category), color = "white", size=4) +
  #scale_fill_brewer(palette="Set1")
```
:::

::: {.column width="50%"}

- Avoid using pie charts;

- When you have many slices and/or the slices are roughly the same size, it becomes impossible to read. 

- In general, Bar Charts are much easier to read than Pie Charts -- so stick with them. 
  
:::

::::

# Two categorical Variables

## Contigency tables

-   Appropriate for summarizing the counts of two categorical variables;

-   Facilitates the analysis of the relationship between two categorical variables;

## Case Study: Aspirin and Heart Attack

-   A controlled, randomized, double-blind study on the effects of aspirin was conducted in the 80s.

::: two-way-tbl
| Drug    | Heart Attack | No Heart Attack | Total  |
|---------|--------------|-----------------|--------|
| Aspirin | 104          | 10,933          | 11,037 |
| Placebo | 189          | 10,845          | 11,034 |
| Total   | 293          | 21,778          | 22,071 |

: Data from the Aspirin Study {#tbl-aspirin-study}
:::

::: {.fragment .fade-in style="font-size:0.75em;"}
*The study was reported on the front page of the New York Times on January 27, 1988. (See the article [here](https://www.nytimes.com/1988/01/27/us/heart-attack-risk-found-to-be-cut-by-taking-aspirin.html).)*
:::

## A few questions of interest

:::{style='font-size:0.9em;'}
<ul>
  <li style='margin:0.75em 0 !important;'> What proportion of individuals did not have a heart attack? </li>
  <li style='margin:0.75em 0 !important;'> What proportion of individuals received a placebo? </li>
  <li style='margin:0.75em 0 !important;'> What proportion of individuals had a heart attack were on aspirin? </li>
  <li style='margin:0.75em 0 !important;'> What proportion of individuals who were on aspirin had a heart attack? </li>
  <li style='margin:0.75em 0 !important;'> What proportion of individuals who were on placebo and had a heart attack? </li>
</ul>
:::

## Marginal Distribution

-   What is the proportion of individuals who did not have a heart attack?

::: {.fragment .fade-in-then-out .absolute .two-way-tbl style='margin-left: 20vh;'}
| Drug    | Heart Attack | No Heart Attack | Total  |
|---------|--------------|-----------------|--------|
| Aspirin | 104          | 10,933          | 11,037 |
| Placebo | 189          | 10,845          | 11,034 |
| Total   | 293          | 21,778          | 22,071 |

: Data from the Aspirin Study
:::

:::{.fragment style='margin-left: 8vh;'}
| Heart Attack | No Heart Attack |
|--------------|-----------------|
| 293 (1.33%)  | 21,778 (98.67%) |

: Marginal distribution of the occurrence of heart attack {#tbl-marginal-dist-heart}
:::

::: {.fragment style="margin-top: 1em;"}
Answer: 98.67%
:::

## Marginal Distribution

-   What proportion of individuals received placebo?

::: {.fragment .fade-in-then-out .absolute .two-way-tbl  style='margin-left: 20vh;'}
| Drug    | Heart Attack | No Heart Attack | Total  |
|---------|--------------|-----------------|--------|
| Aspirin | 104          | 10,933          | 11,037 |
| Placebo | 189          | 10,845          | 11,034 |
| Total   | 293          | 21,778          | 22,071 |

: Data from the Aspirin Study
:::

::: {.fragment style='margin-left: 8vh;'}
| Aspirin      | Placebo      |
|--------------|--------------|
| 11,037 (50%) | 11,034 (50%) |

: Marginal distribution of drug type {#tbl-marginal-dist-drug}
:::

::: {.fragment style="margin-top: 1em;"}
Answer: 50%
:::

## Conditional Distribution

-   What proportion of individuals who had a heart attack were on aspirin?

::: {.fragment .fade-in-then-out .absolute .two-way-tbl  style='margin-left: 20vh;'}
| Drug    | Heart Attack | No Heart Attack | Total  |
|---------|--------------|-----------------|--------|
| Aspirin | 104          | 10,933          | 11,037 |
| Placebo | 189          | 10,845          | 11,034 |
| Total   | 293          | 21,778          | 22,071 |

: Data from the Aspirin Study
:::

::: {.fragment style="margin-left: 8vh;"}
| Aspirin      | Placebo      |
|--------------|--------------|
| 104 (35.49%) | 189 (64.51%) |

: Conditional distribution of drug type among people who had a heart attack {#tbl-conditional-dist-drug}
:::

::: {.fragment .absolute .fade-in-then-out style="margin-top: 1em;"}
Answer: 35.49%
:::

::: {.fragment style="margin-top: 1em;"}
*Note that we fix the condition of `occurrence of heart attack` and look at the distribution of `drug type`.*
:::

## Conditional Distribution

-   What proportion of individuals who were on aspirin had a heart attack?

::: {.fragment .fade-in-then-out .absolute .two-way-tbl  style='margin-left: 20vh;'}
| Drug    | Heart Attack | No Heart Attack | Total  |
|---------|--------------|-----------------|--------|
| Aspirin | 104          | 10,933          | 11,037 |
| Placebo | 189          | 10,845          | 11,034 |
| Total   | 293          | 21,778          | 22,071 |

: Data from the Aspirin Study
:::

::: {.fragment style="margin-left: 8vh;"}
| heart attack | no heart attack |
|--------------|-----------------|
| 104 (0.94%)  | 10,933 (99.06%) |

: Conditional distribution of the occurrence of heart attack among people who were on `aspirin` {#tbl-conditional-dist-heart}
:::

::: {.fragment .absolute .fade-in-then-out style="margin-top: 1em;"}
Answer: 0.94%
:::

::: {.fragment style="margin-top: 1em;"}
*Note that we fix the condition of `drug type` and look at the distribution of `occurrence of heart attack`.*
:::

## Intersection

-   What proportion of individuals who were on placebo and had a heart attack?

::: {.fragment .two-way-tbl}
| Drug    | Heart Attack | No Heart Attack | Total  |
|---------|--------------|-----------------|--------|
| Aspirin | 104          | 10,933          | 11,037 |
| Placebo | 189          | 10,845          | 11,034 |
| Total   | 293          | 21,778          | 22,071 |

: Data from the Aspirin Study
:::

::: {.fragment style="margin-top: 1em;"}
Answer: $\frac{189}{22071} = 0.86\%$
:::

## Association

-   We can use the contingency table to explore the relationship between the variables.

::: {.fragment .fade-in-then-out .absolute}
-   What would you expect to see if there was no association between having a heart attack and the drug taken?
:::
::: {.fragment .fade-in-then-out .absolute}

-   What would you expect to see if there was no association between having a heart attack and the drug taken?
    -   [roughly the same proportion of `heart attacks` in the `placebo` and `aspirin` groups;]{style='font-size:0.9em;'}
:::
::: {.fragment .fade-in-then-out .absolute}

-   What would you expect to see if there was no association between having a heart attack and the drug taken?
    -   [the proportion in the marginal distribution should match the one in the conditional distribution;]{style='font-size:0.9em;'}
:::


## Association

::: {.two-way-tbl style='margin-bottom:1em;'}

| Drug Type | Heart Attack | No Heart Attack | Total  |
|-----------|--------------|-----------------|--------|
|  Aspirin  |     104      |      10,933     | 11,037 |
|  Placebo  |     189      |      10,845     | 11,034 |
|  Total    |     293      |      21,778     | 22,071 |

: Data from the Aspirin Study
:::

:::{.fragment .absolute .fade-in-then-out style='font-size:0.75em;'}
- Proportion of people who had a heart attack:
- Proportion of people on `aspirin` who had a heart attack: 
- Proportion of people on `placebo` who had a heart attack:
:::

:::{.fragment .absolute .fade-in-then-out style='font-size:0.75em;'}
- Proportion of people who had a heart attack: $\frac{293}{22071}=$ 1.33%
- Proportion of people on `aspirin` who had a heart attack: $\frac{104}{11037}=$ 0.94%
- Proportion of people on `placebo` who had a heart attack: $\frac{189}{11034}=$ 1.71%
:::

## Activity {.scrollable}

<div style='font-size:0.6em; margin-bottom: 1em;'>
Consider two over-the-counter topical ointment brands (`Brand A` and `Brand B`) for treating eczema.  Eczema patients who have used either brand were asked about the effectiveness of the ointments (whether or not the eczema is cured within two weeks).  The results are summarized in the table below:
</div>

<table id='half_slide_table'>
<caption style='font-size: 1em; width: 50vh;'>Table 6: Results of Eczema trial.</caption>
<thead>
  <tr>
    <th></th>
    <th colspan="2">Eczema</th>
    <th></th>
  </tr>
  <tr>
    <th>Ointment Brand</th>
    <th>Cured</th>
    <th>Not Cured</th>
    <th>Total</th>
  </tr>
</thead>
<tbody>
  <tr>
    <td>Brand A</td>
    <td>100 (40%)</td>
    <td>150 (60%)</td>
    <td>250 (100%)</td>
  </tr>
  <tr>
    <td>Brand B</td>
    <td id='independent-cell'>200 (50%)</td>
    <td id='dependent-cell'>200 (50%)</td>
    <td>400 (100%)</td>
  </tr>
  <tr>
    <td>Total</td>
    <td id='total-column-independent'>300 (46%)</td>
    <td id='total-column-dependent'>350 (54%)</td>
    <td>650 (100%)</td>
  </tr>
</tbody>
</table>

<div style='width: 60%; margin-left: auto; margin-right: auto;'>

```{ojs}
//| id: test

viewof count = {
  let input = Inputs.range([0, 400], 
                           {value: 200,
                            step: 1, 
                            label: "# patients cured after using Brand B"});
  d3.select(input).select('input[type="number"]').style("display", "none");
  d3.select(input).style("font-size", "0.55em").style('width', '100%');
  d3.select(input).select("label").style('width', '35%');
  //d3.select("#test").style("width", '35%');
  return input;
}

update_entry = (view) => {
  let cell_independent = document.querySelector('#independent-cell');
  let cell_dependent = document.querySelector('#dependent-cell');
  let cell_total_independent = document.querySelector('#total-column-independent');
  let cell_total_dependent = document.querySelector('#total-column-dependent');
  cell_independent.textContent = `${view} (${(100 * view / 400).toFixed(1)}%)`;
  cell_dependent.textContent = `${400 - view} (${(100*(400 - view)/400).toFixed(1)}%)`;
  cell_total_independent.textContent = `${100 + view} (${(100*(100 + view)/650).toFixed(1)}%)`;
  cell_total_dependent.textContent = `${150 + (400 - view)} (${(100*(150 + (400 - view))/650).toFixed(1)}%)`;
}
```

</div>
```{ojs}
//| output: false
update_entry(count)
```

:::: {.columns style="display: flex; justify-content: center; width: 1000px; align-items: center;"}

::: {.column width="40%"}
<div style='font-size: .55em; margin-top: 2em;'>**Presence of `Eczema` among people who used `Brand` `A` ointment**</div>

```{ojs}
data = [
  {"X": "Brand A", "Eczema": "Cured",     "Frequency": 100, "Proportion": 100 / (100 + 150)},
  {"X": "Brand A", "Eczema": "Not cured", "Frequency": 150, "Proportion": 150 / (100 + 150)},
  {"X": "Brand B", "Eczema": "Cured",     "Frequency": count, "Proportion": count / 400},
  {"X": "Brand B", "Eczema": "Not cured", "Frequency": 200, "Proportion": (400 - count) / 400}]


barPlot = Plot.plot({
  x: {
    //axis: null,
    domain: data.Y,
    //range: [150, 300]
  },
  y: {
    grid: true,
    domain: [0, 1]
  },
  marks: [
    Plot.barY(data.filter(d => d.X == 'Brand A'), {x: "Eczema", y: "Proportion", fill: "X", title: "X"}),
    Plot.ruleY([0])
  ],
  style: {
    fontSize: '.45em',
  },
  height: 290,
	width: 478,
	marginLeft: 50,
	marginRight: 50,
	marginTop: 50,
	marginBottom: 50,
});
```
:::


::: {.column width="40%"}

<div style='font-size: .55em; margin-top: 2em;'>**Presence of `Eczema` among people who used `Brand` `B` ointment**</div>

```{ojs}
Plot.plot({
  x: {
    domain: data.Y,
  },
  y: {
    grid: true,
    domain: [0, 1]
  },
  marks: [
    Plot.barY(data.filter(d => d.X == 'Brand B'), {x: "Eczema", y: "Proportion", fill: "X", title: "X"}),
    Plot.ruleY([0])
  ],
  style: {
    fontSize: '.45em',
  },
  height: 290,
	width: 478,
	marginLeft: 50,
	marginRight: 50,
	marginTop: 50,
	marginBottom: 50,
});

```
:::
::::

<div style='height: 100px;'>
</div>

# Case study

## Race and Death Penalty in Florida

- Does race affect the chance of receiving a death penalty sentence in Florida?

- @radelet_1981 examined data on homicide indictments in 20 Florida counties between 1976 and 1977.

------------------------------------------------------------------------

:::{.two-way-tbl }
<table style='width: 80vh; margin-top:3em;'>
<caption>Table 6: Death Penalty sentences from homicide indictments between 1976 and 1977 in Florida [@radelet_1981] </caption>
<thead>
  <tr>
    <th></th>
    <th colspan="2">Death Penalty</th>
    <th></th>
  </tr>
  <tr>
    <th>Race</th>
    <th>Yes</th>
    <th>No</th>
    <th>Total</th>
  </tr>
</thead>
<tbody>
  <tr>
    <td>White</td>
    <td>19</td>
    <td>141</td>
    <td>160</td>
  </tr>
  <tr>
    <td>Black</td>
    <td>17</td>
    <td>149</td>
    <td>166</td>
  </tr>
  <tr>
    <td>Total</td>
    <td>36</td>
    <td>290</td>
    <td>326</td>
  </tr>
</tbody>
</table>
:::


## Case Study: Questions of interest
-   <u>**Question 1:**</u> What proportion of indictments resulted in a death sentence?
-   <u>**Question 2:**</u> Among those who were sentenced to death, what is the proportion of white defendants?
-   <u>**Question 3:**</u> Compare the proportion of white defendants sentenced to death against that of black defendants.

## Case Study: Diving Deeper


:::{#tbl-panel layout-ncol=2 style='margin-top:2em;'}
:::{.two-way-tbl}
<table style='min-width: 40vh;'>
<caption>Table 7: White victims [@radelet_1981] </caption>
<thead>
  <tr>
    <th></th>
    <th colspan="2">Death Penalty</th>
    <th></th>
  </tr>
  <tr>
    <th>Race</th>
    <th>Yes</th>
    <th>No</th>
    <th>Total</th>
  </tr>
</thead>
<tbody>
  <tr>
    <td>White</td>
    <td>19</td>
    <td>132</td>
    <td>151</td>
  </tr>
  <tr>
    <td>Black</td>
    <td>11</td>
    <td>52</td>
    <td>63</td>
  </tr>
  <tr>
    <td>Total</td>
    <td>30</td>
    <td>184</td>
    <td>214</td>
  </tr>
</tbody>
</table>
:::

:::{.two-way-tbl }
<table style='min-width: 40vh;'>
<caption>Table 8: Black victims [@radelet_1981] </caption>
<thead>
  <tr>
    <th></th>
    <th colspan="2">Death Penalty</th>
    <th></th>
  </tr>
  <tr>
    <th>Race</th>
    <th>Yes</th>
    <th>No</th>
    <th>Total</th>
  </tr>
</thead>
<tbody>
  <tr>
    <td>White</td>
    <td>0</td>
    <td>9</td>
    <td>9</td>
  </tr>
  <tr>
    <td>Black</td>
    <td>6</td>
    <td>97</td>
    <td>103</td>
  </tr>
  <tr>
    <td>Total</td>
    <td>6</td>
    <td>106</td>
    <td>112</td>
  </tr>
</tbody>
</table>
:::

Death Penalty sentences grouped per victims' races
:::

## Case Study: Questions of interest
-   <u>**Question 4:**</u> Compare the proportions of indictments that resulted in a death sentence for black and white victims.
-   <u>**Question 5:**</u> For black victims, compare the proportion of white defendants who were sentenced to death against black defendants.
-   <u>**Question 6:**</u> For white victims, compare the proportion of white defendants who were sentenced to death against black defendants.

## Simpson's Paradox
- <u>**Question 7:**</u> Compare your results in Questions 6 and 7 with the result you obtained in Question 3. Is it surprising?

:::{.fragment}
- This reversal in the direction of association when accounting for a third variable is called *Simpson's Paradox*; 
:::

## References

